---
title: "Course DFW Trends"
author: |
    | Joel Schwartz
    | University Initiatives and Student Success
    | California State University, Sacramento
date: "April 14, 2022"
output:
  xaringan::moon_reader:
    #css: [default, "Wilke-slides-theme.css"]
    #chakra: libs/remark-latest.min.js
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      #slideNumberFormat: ''
      titleSlideClass: [center, middle]
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, 
                      #fig.retina=3, # Poor quality, still didn't display in Chrome
                      dev="svg",
                      fig.asp=0.618, fig.width=9)

library(splines)
library(ggh4x)
library(knitr)
library(kableExtra)
library(colorspace)
library(udwverse)
library(ggdist)
library(ggstance)

theme_set(theme_classic2(cap.size=0.55) + 
            theme(strip.text=element_text(size=rel(0.9)),
                  panel.background = element_rect(colour = "grey65", 
        fill = NA, size = 0.25)))

#theme_set(theme_bw())

con=connect_to_udw()

```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r load-premade-data}

cls = read_fst("cls.for.dfw.fst")
cls = cls %>% mutate(urm.ipeds.term = get_urm_status(eth.ipeds.term))
cls = cls %>% filter(term < current_term())


l = read_fst("l.for.dfw.fst")

```


```{r helper-functions}
source("r/functions.r")

```


### Course DFW rate, all undergraduates

.small-font[
```{r p-dfw-trend, eval=FALSE}
pdat = cls %>%
  filter(erss=="Y", grepl("Fall|Spring", term), 
  ) %>% 
  summarize_all_group_combinations(
    value.vars=dfw.flg,
    static.groups=term.grps,
    combo.groups = c("plan.college1", term.demog.grps),
    FUNS = funs.wt,
    max.k=1
  ) %>% 
  clean_up() %>% 
  mutate(Group=abbreviate_labels(Group, "max")) %>% 
  filter(term.abbr >= min.term.abbr)

plot_fnc(pdat, term.abbr, dfw.flg_mean, ylim=c(0, 0.26), caption="Term")
```
]

.center[
```{r p-dfw-trend-out, ref.label="p-dfw-trend"}
```
]

---

### Course DFW rate, first-time students in first semester

.small-font[
```{r p-dfwftf-trend, eval=FALSE}
pdat = cls %>%
  filter(erss=="Y", grepl("Fall|Spring", term)) %>% 
  summarize_all_group_combinations(
    value.vars=dfw.flg,
    static.groups=c(term.grps, coh.grps),
    combo.groups = c("plan.college1", term.demog.grps),
    FUNS = funs.wt,
    max.k=1
  ) %>% 
  clean_up() %>% 
  mutate(Group=abbreviate_labels(Group, "max")) %>% 
  filter(coh.term.abbr >= min.coh.term.abbr)

pdat %>% 
  filter(grepl("F", coh.term.abbr), grepl("Fr", coh), base.time==1) %>% 
  plot_fnc(coh.term.abbr, dfw.flg_mean, ylim=c(0, 0.35), caption="Cohort Entering Term")
```
]

.center[
```{r p-dfwftf-trend-out, ref.label="p-dfwftf-trend"}
```
]

---

### Courses with DFW rate of greater than 20% for Spring/Fall 2021

.small-font[
```{r p-high-dfw-courses, eval=FALSE}
pdat = cls %>% 
  filter(term %in% c("Spring 2021", "Fall 2021")) %>% #filter(course=="HRS 114") %>% select(emplid, course, urm.ipeds.term, grade)
  filter(unt.taken>=2, erss=="Y", session.code==1)

pdat = pdat %>% 
  summarize_all_group_combinations(dfw.flg, 
                                   static.groups=c("acad.group", "div", "course"),
                                   combo.groups=c("eth.ipeds.foreign.term", "urm.ipeds.term", "gender", 
                                                  "pell.term.flg", "first.gen.no.college.attendance"), 
                                   max.k=1)

# Get overall DFW rates
pdat = pdat %>% 
  arrange(desc(Variable=="All"), desc(dfw.flg_mean)) %>% 
  mutate(course=fct_inorder(course)) %>% 
  arrange(course, Variable, Group)


pdat = pdat %>% mutate(Variable=recode(Variable, 
                                       "IPEDS Race/Ethnicity Category"="Ethnicity",
                                       "First Generation Status (neither parent attended college at all)"="First Generation Status"))


pfnc = function(data, div1, xsize=0.9, n.courses=2000) {
  data %>% 
    filter(div==div1) %>% 
    filter(Group %in% c("All", "URM", "Non-URM")) %>% 
    arrange(desc(Group=="All"), desc(dfw.flg_mean)) %>% 
    mutate(course=fct_inorder(as.character(course))) %>% 
    group_by(acad.group) %>% 
    filter(course %in% unique(course)[1:n.courses]) %>% 
    ggplot(aes(course, dfw.flg_mean, colour=Group, group=course)) + 
      geom_line(colour="grey70", size=0.2) +
      geom_point(aes(size=n)) +
      theme_classic2(base.size=10) +
      theme(panel.grid.major.y=element_line(),
            legend.text=element_text(margin=margin(l=-7)),
            axis.text.x=element_text(size=rel(xsize))) +
            thm_rotate_xvar() +
      scale_y_pct(expd=c(0.01,0.01), limits=c(0,1), breaks=seq(0,1,0.1)) +
      theme(legend.position="bottom") +
      labs_remove() + labs(size="Enrollment") +
      facet_grid(cols=vars(acad.group), scales="free_x", space="free_x") +
      scale_size_continuous(range=c(0.05,5), limits=c(1,2000), breaks=c(20,100,500,1000,1500)) +
        guides(colour=guide_legend(override.aes=list(size=3))) +
    scale_colour_manual(values=c("All"="grey40", hcl(c(0,220), 50, 70) %>%
                                   set_names("URM","Non-URM")))
}


pdat1 = pdat  %>% 
  group_by(course) %>% 
  filter(any(n >= 20), any(Group=="All" & dfw.flg_mean >= .2)) %>% 
  ungroup()

pdat1 %>% 
  pfnc(div1="Upper") + labs(title="Upper Division") +
  pdat1 %>% pfnc(div1="Lower", n.courses=40) + labs(title="Lower Division") +
  plot_layout(ncol=1, guides="collect") +
  plot_annotation(theme=theme(legend.position="bottom", 
                              legend.box.margin=margin(t=-5)))


# cls %>%
#   #filter(term %in% c("Spring 2021", "Fall 2021")) %>%
#   filter(course=="PHYS 135") %>% npct(term, dfw.flg) %>% 
#   filter(dfw.flg)

#pdat %>% filter(course=="HRS 114", Variable %in% c("All", "URM Status")) %>% ha
```
]

.center[
```{r p-high-dfw-courses-out, ref.label="p-high-dfw-courses", fig.width=13, fig.asp=0.52}
```
]


---

### DFW rate and URM equity gap by college and course

```{r}
library(ggh4x)

min.enrl=20

pdat = cls %>%
  filter(term %in% c("Spring 2021", "Fall 2021")) %>% #filter(course=="HRS 114") %>% select(emplid, course, urm.ipeds.term, grade)
  filter(unt.taken>=2, erss=="Y", session.code==1) %>% 
  group_by(course) %>% 
  filter(max(n())>=min.enrl) %>% 
  ungroup

pdat = pdat %>%
  summarize_all_group_combinations(dfw.flg,
                                   static.groups=c("acad.group", "dept","dept.abbr", "div", "course"),
                                   combo.groups=c("urm.ipeds.term"),
                                   max.k=1)

# Get overall DFW rates
pdat = pdat %>%
  arrange(desc(Variable=="All"), div, desc(dfw.flg_mean)) %>%
  mutate(course=fct_inorder(course)) %>%
  arrange(course, Variable, Group)

pdat = pdat %>% 
  group_by(acad.group, dept) %>% 
  mutate(dept = case_when(length(unique(course)) <= 5 ~ dept.abbr,
                          nchar(dept) > 25 & length(unique(course)) <= 10 ~ dept.abbr,
                          TRUE ~ dept)) %>% 
  ungroup

# Get each subject's percent for each college
dept.pct = pdat %>%
  filter(Variable=="All") %>%
  group_by(acad.group, dept) %>%
  tally %>%
  mutate(pct=cumsum(n/sum(n)))


# Get indices of subjects that go in each row
row.max=140
colleges = sort(unique(pdat$acad.group)) %>% set_names()
plot.rows = colleges %>% map(
  ~{
    pdat1 = pdat %>% filter(acad.group==.x)
    dept.pct1 = dept.pct %>% filter(acad.group==.x)

    ntot = sum(dept.pct1$n)

    num.rows = max(round(ntot/ifelse(.x %in% c("ECS","NSM","HHS","SSIS"), 100, row.max)), 1)

    # Cum pct break points
  pct.breaks = seq(0,1, 1/num.rows)
  pct.breaks = pct.breaks[-1]

  breaks = map_int(1:length(pct.breaks),
                   ~which.min(abs(dept.pct1$pct - .x/num.rows)))

  map2(c(0,breaks[-length(breaks)]), breaks,
       ~(.x+1):.y)
  }
)

tabs = list()

maxy = round(max(pdat$dfw.flg_mean)*10)/10

for(i in colleges) {

  
  pl = map(plot.rows[[i]],
           ~{
             
             pdept = dept.pct %>%
               filter(acad.group==i) %>%
               slice(.x) %>%
               pull(dept)
             
             pdat = pdat %>%
               filter(dept %in% pdept) 
             
             # Place lower/upper separator
             x.vline = pdat %>% 
               filter(Variable=="All", div=="Lower") %>%
               count(dept) %>% 
               ungroup %>% 
               mutate(n = n + 0.5)
             
             pdat %>%
               ggplot(aes(course, dfw.flg_mean, colour=Group, group=course)) +
               geom_line(colour="grey70", size=0.2) +
               geom_point(aes(size=n)) +
               geom_vline(data=x.vline, aes(xintercept=n),
                          size=0.25, colour="grey85") +
               theme_classic2(base.size=10) +
               theme(panel.grid.major.y=element_line(),
                     strip.text.x=element_text(size=rel(1.05)),
                     legend.text=element_text(margin=margin(l=-7))) +
               thm_rotate_xvar() +
               scale_y_pct(expd=c(0.01,0.01), limits=c(0,1), breaks=seq(0,1,0.1)) +
               theme(legend.position="bottom",
                     axis.text.x=element_text(size=rel(ifelse(.x=="ALS", 0.7, 0.8))),
                     axis.text.y=element_text(siz=rel(0.7))) +
               labs_remove() + labs(size="Enrollment") +
               facet_grid(cols=vars(dept), scales="free_x", space="free_x") +
               scale_size_continuous(range=c(0.05,5), limits=c(1,2000), breaks=c(20,100,500,1000,1500)) +
               guides(colour=guide_legend(override.aes=list(size=3))) +
               scale_colour_manual(values=c("All"="grey40", hcl(c(0,220), 50, 70) %>%
                                              set_names("URM","Non-URM"))) +
               coord_cartesian(ylim=c(0,maxy))
             
           })
  
  p = wrap_plots(pl, ncol=1, guides="collect") +
    plot_annotation(theme=theme(legend.position="bottom",
                                legend.box.margin=margin(t=-7,b=-12)),
                    caption=paste("Data shown for courses worth 2+ units and with 20+ total enrollment for Spring/Fall 2021"))
  
  tabs[[i]] = p
}


```



.panelset[

.panel[.panel-name[ALS]

```{r pchunk1, fig.width=13, fig.asp=0.45}

tabs[["ALS"]]

```
]


.panel[.panel-name[COB]

```{r pchunk2, fig.width=13, fig.asp=0.45}

tabs[["CBA"]]

```
]

.panel[.panel-name[ECS]

```{r pchunk3, fig.width=13, fig.asp=0.45}

tabs[["ECS"]]

```
]

.panel[.panel-name[ED]

```{r pchunk4, fig.width=13, fig.asp=0.45}

tabs[["ED"]]

```
]

.panel[.panel-name[HHS]

```{r pchunk5, fig.width=13, fig.asp=0.45}

tabs[["HHS"]]

```
]

.panel[.panel-name[NSM]

```{r pchunk6, fig.width=13, fig.asp=0.45}

tabs[["NSM"]]

```
]

.panel[.panel-name[SP]

```{r pchunk7, fig.width=13, fig.asp=0.45}

tabs[["SP"]]

```
]

.panel[.panel-name[SSIS]

```{r pchunk8, fig.width=13, fig.asp=0.45}

tabs[["SSIS"]]

```
]

]



---

### First-semester GPA and student academic and socioeconomic background 

.small-font[
```{r p-grades-vs-ses, eval=FALSE}

pfnc = function(data, yvar, ylab=NULL, min.coh.term.abbr="F17", points=TRUE, 
                .se=FALSE, mytitle=NULL) {
  data = data %>% 
    filter(coh.term.abbr>= min.coh.term.abbr, 
           grepl("Fall", coh.term)) %>% 
    mutate(period = case_when(coh.term.abbr <= "F19" ~ paste0(min.coh.term.abbr,"-F19"),
                              coh.term.abbr >= "F20" ~ as.character(coh.term.abbr)),
           period=factor(period, levels=c(paste0(min.coh.term.abbr, "-F19"), paste0("F", 20:30)))) %>% 
    select(emplid, period, {{yvar}}, 
           "HS GPA"=hs.gpa, "SAT Math"=sat.math, "SAT Verbal"=sat.verbal, 
           "11th Grade Math"=EAPMS, "11th Grade English"=EAPES,
           "HS Free/Red Price Meal %"=last.school.attended.frpm.pct,
           "Zip Code % HS degree or less"=zip.pct.high.school.or.less,
           "Zip Code Median Income (thousands)"=home.zip.median.income,
           "Adjusted Gross Income (thousands)"=agi.parents) %>% 
    pivot_longer(-c(emplid, period, {{yvar}})) %>% 
    filter(!(name=="Adjusted Gross Income (thousands)" & !between(value, -1e4,3e5)),
           !(name=="Zip Code Median Income (thousands)" & !between(value, 25000, 1.5e5))) %>% 
    group_by(name, period) %>% 
    filter(mean(!is.na(value)) > 0.5) %>% 
    ungroup %>% 
    mutate(name=factor(name, levels=unique(name))) %>% 
    group_by(name) %>% 
    group_split() 
  
  #print(map(data, ~.x %>% count(name, is.na(value))))
  
  p = data %>% 
    map(
      ~{ p = .x %>% 
        ggplot(aes(value, {{yvar}}, colour=period))
      
      if(points) {
        p = p +
          geom_point(size=0.3, colour="grey60", alpha=0.05) 
      }
      
      p = p +
        geom_smooth(se=.se, size=0.7, 
                    #method="loess", span=1.5,
                    method="gam", formula=y ~ s(x, bs="cs", k=3)) +
        facet_wrap(vars(name), scales="free_x", nrow=1) +
            labs_remove() + labs(y=ylab) +
        guides(colour=guide_legend(nrow=1)) +
        scale_colour_manual(values=c(viridis_pal(end=0.7)(2), "red") %>% 
                              set_names("F17-F19", "F20","F21")) +
        theme_classic2(base.size=10)
      
      rngy = .x %>% pull({{yvar}})
      rngy = range(rngy, na.rm=TRUE)
      if(rngy[2] <= 1) {
        p = p + scale_y_pct()
      } else {
        p = p + scale_y_grade()
      }
      
      if(grepl("Income", .x$name[1])) {
        p = p + scale_x_continuous(labels=dollar_format(scale=1e-3), 
                                   breaks=pretty_breaks(3))
      } else if(grepl("GPA|SAT|11th", .x$name[1])) {
        
        lim = quantile(.x$value, na.rm=TRUE, probs=c(0.005, 0.995))
        
        p = p + scale_x_continuous( breaks=pretty_breaks(3)) + 
          coord_cartesian(xlim=lim)
      } else { 
        p = p + scale_x_pct()
      }
      p
      }
    )
  
  #p[1:8] = map(p[1:8], ~.x + guides(colour="none"))
  p[-c(1,6)] = map(p[-c(1,6)], ~.x + labs(y=NULL) + 
                theme(axis.ticks.y=element_blank(),
                      axis.text.y=element_blank()))
  
  wrap_plots(p, ncol=5, guides="collect") +
    plot_annotation(theme=theme(legend.position="bottom"),
                    subtitle=mytitle)
}


p1 = l %>% 
  filter(base.time==1, grepl("Fr", coh)) %>% 
  pfnc(ssr.cum.en.gpa, ylab="First-semester GPA")
#p2 = l %>% filter(plan.college1st == college) %>% pfnc(mytitle = college)

p1
```

```{r eval=FALSE}
# p1 = l %>% 
#   filter(base.time==1, grepl("Fr", coh), plan.college1=="ECS") %>% 
#   pfnc(ssr.cum.en.gpa)
# #p2 = l %>% filter(plan.college1st == college) %>% pfnc(mytitle = college)
# 
# p1

p2 = l %>%
  filter(base.time==8,  grepl("Fr", coh), plan.college1st=="ECS") %>% 
  mutate(grad4 = as.numeric(pst.flgERS=="Graduated")) %>% 
  pfnc(grad4, min.coh.term.abbr="F16", mytitle="Probability of graduating in four years for students who start as Engineering majors (Fall 2016/17 first-time cohorts)", points=FALSE) +plot_annotation(theme=theme(legend.position="none"))

p2a = l %>%
  filter(base.time==9, grepl("Fr", coh), plan.college1st=="ECS") %>% 
  mutate(grad4 = as.numeric(pst.flgERS=="Graduated")) %>% 
  pfnc(grad4, min.coh.term.abbr="F17", mytitle="Probability of graduating in 4.5 years for students who start as Engineering majors (Fall 2016/17 first-time cohorts)", points=FALSE) + plot_annotation(theme=theme(legend.position="none"))


p3 = l %>%
  filter(base.time==10, grepl("Fr", coh), plan.college1st=="ECS") %>% 
  mutate(grad4 = as.numeric(pst.flgERS=="Graduated")) %>% 
  pfnc(grad4, min.coh.term.abbr="F16", mytitle="Probability of graduating in five years for students who start as Engineering majors (Fall 2016 first-time cohort)", points=FALSE) + plot_annotation(theme=theme(legend.position="none"))


pdat = l %>% 
  filter(full.time.1st.flg!="Part time") %>% 
  convert_outcomes_to_logical() %>% 
  summarize_all_group_combinations(pst.flgERSproj, 
                                   static.groups=c(coh.grps, "plan.college1st"), 
                                   combo.groups=c("eth.ipeds.foreign"),
                                   max.k=2) 
pdat = pdat %>% 
  bind_rows(
    l %>% 
      filter(full.time.1st.flg!="Part time") %>% 
  convert_outcomes_to_logical() %>% 
  summarize_all_group_combinations(pst.flgERSproj, 
                                   static.groups=c(coh.grps), 
                                   combo.groups=c("eth.ipeds.foreign"),
                                   max.k=2) %>% 
    mutate(plan.college1st="All")
  )

p4 = pdat %>%
  filter(base.time==8, 
         grepl("All|ECS", plan.college1st),
         #grepl("All|College,", Variable), 
         !grepl("For|Unk|Two", Group),
         grepl("Fr", coh), 
         grepl("Fall", coh.term), coh.term >="Fall 2009") %>% 
  csus.data.analysis::plot_outcomes(coh.term.abbr, pst.flgERSproj_mean, group=Group) +
  facet_wrap(vars(plan.college1st)) +
  scale_y_continuous(expand=expansion(c(0.01,0)), labels=percent_format(accuracy=1),
                     limits=c(0,1)) +
  labs(title="Four-year graduation rates by ethnicity: All students and students who start as Engineering majors",
       x="Cohort entering term",
       caption="Shaded term is a projection")

p4 =p4 + 
  geom_tile(data=tibble(coh.term.abbr="F18", Group=NA), aes(x="F18", y=0.5), fill="grey10", alpha=0.05, colour=NA, width=0.5)

p5 = pdat %>%
  filter(base.time==12, 
         grepl("All|ECS", plan.college1st),
         #grepl("All|College,", Variable), 
         !grepl("For|Unk|Two", Group),
         grepl("Fr", coh), 
         grepl("Fall", coh.term), coh.term >="Fall 2009") %>% 
  csus.data.analysis::plot_outcomes(coh.term.abbr, pst.flgERSproj_mean, group=Group) +
  facet_wrap(vars(plan.college1st)) +
  scale_y_continuous(expand=expansion(c(0.01,0)), labels=percent_format(accuracy=1),
                     limits=c(0,1)) +
    labs(title="Six-year graduation rates by ethnicity: All students and students who start as Engineering majors",
         x="Cohort entering term",
       caption="Shaded term is a projection")

p5 =p5 + 
  geom_tile(data=tibble(coh.term.abbr="F16", Group=NA), aes(x="F16", y=0.5), fill="grey10", alpha=0.05, colour=NA, width=0.5)

# Graduate in Engineering
pdat.grad.same.college = l %>% 
  filter(base.time==12) %>% 
  mutate(deg.plan1 = case_when(term==current_term() & is.na(deg.plan1) & pst.flgERSproj=="Graduated" ~ acad.plan1,
                               TRUE ~ deg.plan1)) %>% 
  filter(full.time.1st.flg!="Part time") %>% 
  left_join(
    get_dept_and_college_by_plan() %>% 
      select(deg.plan1=acad.plan, deg.college1=plan.college)
  ) %>% 
  mutate(pst.flgERSproj = case_when(is.na(deg.college1) ~ pst.flgERSproj,
                                    deg.college1==plan.college1st & pst.flgERSproj=="Graduated" ~ "Graduated",
                                    deg.college1!=plan.college1st & pst.flgERSproj=="Graduated" ~ "other major",
                                    TRUE ~ pst.flgERSproj)) %>% 
  convert_outcomes_to_logical() %>% 
  summarize_all_group_combinations(pst.flgERSproj, 
                                   static.groups=c(coh.grps, "plan.college1st"), 
                                   combo.groups=c("eth.ipeds.foreign"),
                                   max.k=2)

pdat.grad.diff.college = l %>% 
  filter(base.time==12) %>% 
  mutate(deg.plan1 = case_when(term==current_term() & is.na(deg.plan1) & pst.flgERSproj=="Graduated" ~ acad.plan1,
                               TRUE ~ deg.plan1)) %>% 
  filter(full.time.1st.flg!="Part time") %>% 
  left_join(
    get_dept_and_college_by_plan() %>% 
      select(deg.plan1=acad.plan, deg.college1=plan.college)
  ) %>% 
  mutate(pst.flgERSproj = case_when(is.na(deg.college1) ~ pst.flgERSproj,
                                    deg.college1==plan.college1st & pst.flgERSproj=="Graduated" ~ "same major",
                                    deg.college1!=plan.college1st & pst.flgERSproj=="Graduated" ~ "Graduated",
                                    TRUE ~ pst.flgERSproj)) %>% 
  convert_outcomes_to_logical() %>% 
  summarize_all_group_combinations(pst.flgERSproj, 
                                   static.groups=c(coh.grps, "plan.college1st"), 
                                   combo.groups=c("eth.ipeds.foreign"),
                                   max.k=2)

p5a1 = pdat.grad.same.college %>%
  filter(grepl("ECS", plan.college1st),
         #grepl("All|College,", Variable), 
         !grepl("For|Unk|Two", Group),
         grepl("Fr", coh), 
         grepl("Fall", coh.term), coh.term >="Fall 2009") %>% 
  csus.data.analysis::plot_outcomes(coh.term.abbr, pst.flgERSproj_mean, group=Group) +
  facet_wrap(vars(plan.college1st)) +
  scale_y_continuous(expand=expansion(c(0.01,0)), labels=percent_format(accuracy=1),
                     limits=c(0,1)) +
    labs(title="Engineering degree", x="Cohort entering term")

p5a2 = pdat.grad.diff.college %>%
  filter(grepl("ECS", plan.college1st),
         #grepl("All|College,", Variable), 
         !grepl("For|Unk|Two", Group),
         grepl("Fr", coh), 
         grepl("Fall", coh.term), coh.term >="Fall 2009") %>% 
  csus.data.analysis::plot_outcomes(coh.term.abbr, pst.flgERSproj_mean, group=Group) +
  facet_wrap(vars(plan.college1st)) +
  scale_y_continuous(expand=expansion(c(0.01,0)), labels=percent_format(accuracy=1),
                     limits=c(0,1)) +
    labs(title="Other degree", x="Cohort entering term")

p5a1 + p5a2 + plot_annotation(title="Of students who started as Engineering majors: Percent graduating in six years")


pdf("Probability of graduation for engineering majors.pdf", 10*1.5, 5*1.5)
p2 
p2a
p3
dev.off()

pdf("Graduation by ethnicity.pdf", 10, 6)
p4 
p5
list(p5a1, p5a2) %>%  
  map(~.x + theme(strip.text.x=element_blank(), strip.background=element_blank()) +
          geom_tile(data=tibble(coh.term.abbr="F16", Group=NA), aes(x="F16", y=0.5), fill="grey10", alpha=0.05, colour=NA, width=0.5)) %>% 
  wrap_plots() + plot_annotation(title="Of first-time students who started as Engineering majors: Percent graduating in six years",
                                 caption="Shaded term is a projection")
dev.off()






```
]



.center[
```{r p-grades-vs-ses-out, ref.label="p-grades-vs-ses", fig.width=fw, fig.asp=0.5}
```
]

---

### Academic background and socioeconomic status by ethnicity

.small-font[
```{r p-background-by-eth, eval=FALSE}
options(scipen=10)
pdat = l %>% 
  filter(base.time==1, coh.term>="Fall 2018", grepl("Fr", coh), grepl("F", coh.term)) %>%
  mutate(pell.coh.term.flg=pell.coh.term.flg=="Pell",
         first.gen.no.college.attendance=first.gen.no.college.attendance=="First generation") %>% 
  select(eth.ipeds.foreign,  
         #pell.coh.term.flg, first.gen.no.college.attendance,  
      
         hs.gpa, sat.math, sat.verbal, EAPMS, EAPES,
          agi.parents,
         home.zip.median.income, zip.pct.high.school.or.less, 
         last.school.attended.frpm.pct) %>% 
  pivot_longer(-c(eth.ipeds.foreign)) %>% 
  mutate(eth.ipeds.foreign = abbreviate_labels(eth.ipeds.foreign, "max")) %>% 
  mutate(eth.ipeds.foreign = fct_relevel(eth.ipeds.foreign, 
                                         "Afr Amer", "Hispanic","Nat Amer","Pac Isl")) %>% 
  mutate(name = recode(name, !!!get_name_recodes())) %>% 
  mutate(name = factor(name, levels=unique(name))) %>% 
    filter(!grepl("For|Unk", eth.ipeds.foreign))
  
pdat %>% 
  ggplot(aes(eth.ipeds.foreign, value)) + 
  geom_hline(data=. %>% 
               group_by(name) %>%  
               median_qi(value, na.rm=TRUE, .width=0.5) %>% 
               ungroup %>% 
               pivot_longer(c(value, .lower, .upper), values_to="x", names_to="nm"),
             aes(yintercept=x), colour="grey80", size=0.3) + 
  stat_pointinterval(point_interval=median_qi, .width=c(0.5), size=0.5, point_size=1.5,
                     aes(colour=eth.ipeds.foreign)) +
  facet_wrap(vars(name), ncol=5, scales="free_y") +
  thm_rotate_xvar() +
  guides(colour="none") +
  #scale_y_continuous(labels=comma) +
  labs_remove() +
  labs(caption="Points are medians. Ranges are 25th - 75th percentiles. Gray lines are overall medians and 25th - 75th percentiles.") +
  theme(plot.caption=element_text(size=rel(0.65), margin=margin(t=2)),
        strip.text=element_text(size=rel(0.7)))

```
]

.center[
```{r p-background-by-eth-out, ref.label="p-background-by-eth", fig.width=fw, fig.asp=0.45}
```
]


---

### Course grades by instructor

.small-font[
```{r grade-by-instructor, eval=FALSE}
byprof = cls %>%
  filter(unt.taken >= 3, erss=="Y") %>% 
  filter(term %in% most_recent_terms(7)) %>%
  mutate(period = case_when(term <= "Fall 2019" ~ "Pre-pandemic",
                            term=="Spring 2020" ~ "Pandemic begins",
                            term %in% most_recent_terms(3) ~ "During pandemic")) %>% 
  group_by(course, term, class.section) %>% 
  # CHECK THIS: REMOVE SECTIONS WITH MOSTLY NON-UGD
  #filter(! mean(acad.prog %in% c("UGT","UGED")) > 0.2) %>% 
  ungroup %>% 
  summarize_all_group_combinations(dfw.flg,
                                   static.groups=c("period","acad.group", "dept.abbr", "subject",
                                                   "course", "course.descr", "course.numeric",
                                                   "div"),
                                   combo.groups=c("instructor1", "eth.comb.foreign")) %>% 
  mutate(period = fct_relevel(period, "Pre-pandemic", "Pandemic begins"))

min.enrl=200
byprof %>%
  filter(period=="During pandemic", Variable %in% c("All", "instructor1"),
         acad.group %notin% c("SP", "CCE")) %>% 
  ungroup %>% 
  arrange(Group!="All", desc(dfw.flg_mean)) %>% 
  mutate(course=factor(course, levels=unique(course))) %>%
  group_by(course) %>% 
  filter(n() >= 5, any(n >= min.enrl)) %>% #distinct(course)
  arrange(desc(course)) %>% 
  group_by(acad.group) %>%
  arrange(Group != "All", desc(dfw.flg_mean)) %>% 
  filter(course %in% unique(course)[1:10]) %>% #group_by(course) %>% filter(Group=="All")
  ggplot(aes(course, dfw.flg_mean, group=course)) +
    geom_hline(yintercept=0.2, colour="grey80", size=0.25) + 
    geom_line(colour="grey50", size=0.3) +
    
    geom_point(data=. %>% filter(Group != "All"), aes(size=n), 
               colour="grey20", fill="grey60", shape=21) +
    geom_point(data=. %>% filter(Group == "All"), 
               aes(group=course), colour="red", shape=3, size=1.1) +
    facet_grid(cols=vars(acad.group), scales="free_x", space="free_x") +
    scale_y_continuous(labels=percent, expand=expansion(c(0.01,0.03)),
                       limits=c(0,NA)) +
    theme(strip.text.x=element_text(angle=0, hjust=0.5, margin=margin(l=2,r=2, t=2, b=2)),
          axis.text.x=element_text(size=rel(0.7)),
          legend.title=element_text(size=rel(0.7)),
          legend.text=element_text(size=rel(0.6), hjust=1),
          plot.caption=element_text(margin=margin(t=3))) +
    scale_size_continuous(breaks=c(50,100,200,300,400), 
                          range=c(0,3)) +
    thm_rotate_xvar() +
    labs(x=NULL, y=NULL, size="Enrollment", 
         caption=str_wrap(paste0("Each point is one instructor. Red crosses are the overall DFW rate. Data for courses with at least ", min.enrl, " enrollments during Fall 2020 through Fall 2021."), 150))
```
]

.center[
```{r grade-by-instructor-out, ref.label="grade-by-instructor", fig.width=fw, fig.asp=0.5}
```
]

---

### Course grades by high school GPA and instructor

.small-font[
```{r course-grd-by-instr-hsgpa, eval=FALSE}
cls %>% 
  filter(erss=="Y", grepl("Fall|Spring", term)) %>% 
  filter(term %in% most_recent_terms(3)) %>% #filter(course=="PUBH 50")
  group_by(course) %>% 
  filter(n() > 150, length(unique(instructor.emplid1)) >= 4, 
         mean(is.na(hs.gpa)) < 0.2) %>%   
  mutate(ntot=n()) %>% 
  ungroup %>% 
  arrange(desc(ntot)) %>% 
  group_by(acad.group) %>% 
  filter(course %notin% c("GNST 21")) %>% 
  filter(course %in% unique(course)[1:5]) %>% #count(course)
  ungroup %>% 
  plot_course_grades_by_instructor(group.var=instructor1, terms=most_recent_terms(3),
                                   ncol=7, courses=unique(.$course),
                                   ) +
  scale_y_grade() +
  scale_x_continuous(breaks=pretty_breaks(3)) +
  theme(strip.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.8)),
        legend.text=element_text(size=rel(0.7))) + labs(title=NULL)


```
]

.center[
```{r course-grd-by-instr-hsgpa-out, ref.label="course-grd-by-instr-hsgpa", fig.width=fw, fig.asp=0.52}
```
]



---